Demystifying the Combination of Dynamic Slicing and \\ Spectrum-based Fault Localization

### 2019-02-26 paper/ijcai/paper.tex
boxplot

Our implementation of Critical Slicing is based on the Mozilla Lithium
tool\footnote{Lithium details available at {\footnotesize\url{https://github.com/MozillaSecurity/lithium}} (Accessed \today)}. It takes as input a file
and produces as output a simplified version of that file that
satisfies a user-defined oracle. In our case, the oracle is defined
such that the test produces the same failure manifestation as the one
observed with the test execution on the original program. The
Lithium minimization process starts by determining the initial size---in
number of lines---of chunks to delete from the input file. For that,
it chooses the highest power of two number smaller than the file
size. For example, if the file has 1,000 lines, Lithium sets the
initial chunk size to 512 lines. Then, the tool starts a local search looking
for chunks to exclude from the file. If the chunk satisfies the oracle it is removed.
When no more chunks of that given size can be removed, Lithium
divides the chunk size by two and repeats the search. This iterative
process continues until no more lines can be removed.  If $n$ is the
size of the input file and $m$ is the size of the 1-minimal file found
by Lithium, then Lithium usually performs $O(m\cdot\lg(n))$
iterations. Proofs of the algorithm complexity can be found
elsewhere\footnote{Lithium complexity available at \url{https://github.com/MozillaSecurity/lithium/blob/master/src/lithium/docs/algorithm.md} (Accessed \today)}. Our implementation is publicly
available at \textit{$<$anonymized for double-blind review$>$}.

@@

It is worth noting that this is not a problem that afflicts only CS;
\emph{all} purely dynamic slicing techniques manifest this
problem~\cite{Lin:2018:BDE:3238147.3238163}. To mitigate the issue in
CS, it is necessary to make oracles more specific. For that, our
approach was to use as the slicing criterion the stack trace
associated with the test failure. That decision increases the chances
that the sliced program will follow a path to the error similar to
that followed by the original program.
There are faults whose outputs
may slightly differ from one execution to another. For example,
exceptions including memory addresses identifiers, which change on
each execution. For these cases, it was developed a mechanism to compare the
rest of the output ignoring the identifier of the memory address.

@@

In \textbf{RQ3}, we evaluate the impact of combining \ds{} with \sfl{}
to improve fault localization. The first experiment we ran consists of
measuring the \emph{difference of diagnosis cost} between the two
fault localization techniques. This metric has been previously used in
other
studies~\cite{7390282,ang-perez-van-deursen-rui-2017,Pearson:2017:EIF:3097368.3097441,Perez:2018:LQR:3304889.3304927}.
More specifically, we computed $\Delta{}C = C(\textrm{SFL$^{k}$}) -
C(\textrm{\combpar{k}})$, where $C$ denotes the diagnosis cost and is
obtained by computing the mean position in the ranking of all buggy
statements. $\Delta C <0$ means that \combpar{k} performs worse
compared to its baseline. That could happen if dynamic slicing misses
the faulty statement. $\Delta C=0$ means that the faulty statement
remained in the same position. $\Delta C >0$ means that \combpar{k}
outperformed the baseline, \ie{} the mean position of the faulty
statements is smaller in \combpar{k} compared to that of the
baseline. It is worth noting that we assume perfect bug
understanding~\cite{Parnin:2011:ADT:2001420.2001445}, i.e., if the
developer sees the bug in the ranking, he is able to precisely
determine if it is the faulty one.  Figure \ref{fig:diagnosis}
summarizes the performance of \combpar{k} relative to
\sfl{}$^{k}$. The figure shows a histogram over $\Delta{}C$ for
different values of $k$. Note that (1) the number of cases where the
baseline outperforms \combpar{k} is small (\eg{}, \numOutPerformed\ of
the \numFaults\ faults for $k=10$) and that (2) the number of cases
\combpar{k} outperforms the baseline is significative (\eg{}, $117$ of
the \numFaults\ faults for $k=10$).

%% One
%% of the following reasons justifies the baseline outperforming
%% \combpar{k}:\Mar{following sentence is confusing IMO. suggest to
%%   clarify or remove. aren't points two and three equivalent? as for
%%   one, isn't foo an issue both for \combpar{k} and \sfl{}$^{k}$?} (1)
%% a fault of omission, (2) the test has design limitations which \ds~is
%% not capable of handling, or (3) the buggy-class in not in the top $k$
%% of the highest-ranked faulty classes under analysis.

@@

\begin{figure}[h]
	\vspace{-0.3cm}
		\centering
		\includegraphics[width=0.415\textwidth]{figures/performance.pdf}
		\vspace{-0.4cm}
		\caption{Delta Cost of Diagnosis ($\Delta{}C$) per $k$}
		\label{fig:diagnosis}
\end{figure}

\Mar{Sofia, vc. precisa revisar com mais cuidado os dois proximos paragrafos...}

Figure~\ref{fig:boxplot} shows the distributions of $\Delta{}C$ for
different sets of faults as to better understand the performance of
\combpar{k}. 
\Sof{This is not correct}
On the left-hand-side, we show distributions of
$\Delta{}C$, for different values of $k$, considering all cases where
the predicate $\Delta{}C\geq{}0$ holds, \ie{}, we excluded the
\numOutPerformed\ cases where \sfl{}$^{k}$ outperformed \combpar{k}
(see Figure~\ref{fig:diagnosis}). The boxplots on the right-hand-side
of the figure considers all the cases where the predicate
$\Delta{}C>0$ holds, \ie{}, it additionally excludes from the dataset
the cases of faults where the techniques perform equally.\Sof{Needs to be reverted}
\Mar{sofia,
  are you sure there are no zeros in this dataset? The min. in this
  distibution (as per the base of the plot) seems to be 0.}\Sof{Yes, check hangouts} To sum,
although we observed a relatively small number cases where $\Delta{}C$
is negative (\numOutPerformed) and a high number of cases $\Delta{}C$
is zero (\numEquallyPerformed), we observed that there is a
considerable amount of cases where the use of \combpar{k} would be
very beneficial.

\begin{figure}[h]
		\vspace{-0.2cm}
		\centering
		\includegraphics[width=0.49\textwidth]{figures/boxplot.pdf}
		\vspace{-0.6cm}

		\caption{Distributions of $\Delta{}C$. On the left, distribution
        considering all faults but those manifesting faults of
        omission and engineering problems (227 faults). On the right,
        distribution considering all cases where \combpar{k}
        outperformed the baseline (116 faults).}
		\label{fig:boxplot}
		\vspace{-0.2cm}
\end{figure}

In the following, we present statistical analysis of the techniques.
Table~\ref{table:st} reports descriptive statistics about the
distributions of $C$ for the techniques under different values of $k$.
In bold, we present the results of the $227$ faults that do not
comprise the cases that are result of the limitations mentioned before
(test design and fault of omission). The number in parentheses are
relative to the dataset including all data faults; note that results
for this dataset are \emph{not} encouraging.

@@

\begin{table}[h]
	\tiny
	\centering
	\setlength{\tabcolsep}{3pt}
	\begin{tabular}{@{}c|c|c|c|c@{}}
     \toprule
  & \multicolumn{2}{c|}{$k=5$} & \multicolumn{2}{c}{$k=10$} \\
  & SFL  & \comb\       & SFL                & \comb\              \\ \midrule
Mean  & \textbf{590.7} (584.3) & \textbf{573.4} (679.7)   &  \textbf{512.7} (516.4)    & \textbf{506.6} (614.0)   \\ \midrule
Median & \textbf{46.5} (49.4) & \textbf{42.7} (69.3)      & \textbf{54.5} (53.0) & \textbf{40.0} (56.9)\\ \midrule
Variance & \textbf{1257.0} (1223.1) & \textbf{1239.1} (1352.4)  &  \textbf{1171.5} (1148.1) &  \textbf{1167.0} (1294.9) \\ \midrule
Shapiro-Wilk &\makecell{$W$ = 0.52 \\ $p$ = 1.75 x $10^{-24}$} & \makecell{$W$ = 0.51 \\ $p$ = 1.10 x $10^{-24}$} & \makecell{$W$ = 0.48 \\ $p$ = 2.50 x $10^{-25}$} &  \makecell{$W$ = 0.48 \\ $p$ = 2.26 x $10^{-25}$}  \\ \midrule
Friedman & \multicolumn{4}{c}{\makecell{$\chi^{2}$ = 162.61 \\ p-value = 5.00 x $10^{-35}$}} \\
\bottomrule
\end{tabular}
  \caption {Statistical tests for $C$}
  \label{table:st}
\end{table}

@@

\begin{figure}[h]
	\vspace{-1.8cm}
		\includegraphics[width=0.45\textwidth]{figures/heatmap_nemenyi_result.pdf}
		\caption{Siegel post-hoc analysis results}
		\label{fig:performance}
		\vspace{-0.3cm}
\end{figure}

As the results are statistically significant, it is possible to state
that \combpar{k} has a positive effect on the localization of some
types of bugs. There is indeed a problem on using slicing on fault of
omission bugs that needs further research.  Although, this dataset has
more than $50\%$ of fault of omissions, our simple slicing approach
was capable of improving the ranking of $95$ faults for $k=5$ and
$115$ faults for $k=10$. \Mar{Sofia, acho mais apropriado colocar isto
  em conclusions}These are promissing results from what Automated
Program Repair techniques may directly benefit to improve the accuracy
of their own techniques.

>>>>

Our implementation of Critical Slicing is based on the Mozilla Lithium
tool\footnote{Lithium details available at {\footnotesize\url{https://github.com/MozillaSecurity/lithium}} (Accessed \today)}. It takes as input a file
and produces as output a simplified version of that file that
satisfies a user-defined oracle. In our case, the oracle is defined
such that the test produces the same failure manifestation as the one
observed with the test execution on the original program. The
Lithium minimization process starts by determining the initial size---in
number of lines---of chunks to delete from the input file. For that,
it chooses the highest power of two number smaller than the file
size. For example, if the file has $1,000$ lines, Lithium sets the
initial chunk size to $512$ lines. Then, the tool starts a local search looking
for chunks to exclude from the file. If the chunk satisfies the oracle it is removed.
When no more chunks of that given size can be removed, Lithium
divides the chunk size by two and repeats the search. This iterative
process continues until no more lines can be removed.  If $n$ is the
size of the input file and $m$ is the size of the 1-minimal file found
by Lithium, then Lithium usually performs $O(m\cdot\lg(n))$
iterations. Proofs of the algorithm complexity can be found
elsewhere\footnote{Lithium complexity available at \url{https://github.com/MozillaSecurity/lithium/blob/master/src/lithium/docs/algorithm.md} (Accessed \today)}. Our implementation is publicly
available at \textit{$<$anonymized for double-blind review$>$}.

@@

It is worth noting that this is not a problem that afflicts only CS;
\emph{all} purely dynamic slicing techniques manifest this
problem~\cite{Lin:2018:BDE:3238147.3238163}. To mitigate the issue in
CS, it is necessary to make oracles more specific. For that, our
approach was to use as the slicing criterion the stack trace
associated with the test failure. That decision increases the chances
that the sliced program will follow a path to the error similar to
that followed by the original program.
In addition, there are faults whose outputs
may slightly differ from one execution to another. For example,
exceptions including memory address identifiers, which change on
each execution. For these cases, it was developed a mechanism to compare the
rest of the output ignoring the identifier of the memory address.

@@

In \textbf{RQ3}, we evaluate the impact of combining \ds{} with \sfl{}
to improve fault localization. The first experiment we ran consists of
measuring the \emph{difference of diagnosis cost} between the two
fault localization techniques. This metric has been previously used in
other
studies~\cite{7390282,ang-perez-van-deursen-rui-2017,Pearson:2017:EIF:3097368.3097441,Perez:2018:LQR:3304889.3304927}.
More specifically, we computed $\Delta{}C = C(\textrm{SFL$^{k}$}) -
C(\textrm{\combpar{k}})$, where $C$ denotes the diagnosis cost and is
obtained by computing the mean position in the ranking of all buggy
statements. $\Delta C <0$ means that \combpar{k} performs worse
compared to its baseline. That could happen if dynamic slicing misses
the faulty statement. $\Delta C=0$ means that the faulty statement
remained in the same position. $\Delta C >0$ means that \combpar{k}
outperformed the baseline, \ie{} the mean position of the faulty
statements is smaller in \combpar{k} compared to that of the
baseline. It is worth noting that we assume perfect bug
understanding~\cite{Parnin:2011:ADT:2001420.2001445}, i.e., if the
developer sees the bug in the ranking, he is able to precisely
determine if it is the faulty one.  Figure \ref{fig:diagnosis}
summarizes the performance of \combpar{k} relative to
\sfl{}$^{k}$. The figure shows a histogram over $\Delta{}C$ for
different values of $k$. Note that (1) the number of cases where the
baseline outperforms \combpar{k} is small (\eg{}, \numOutPerformed\ of
the \numFaults\ faults for $k=10$) and that (2) the number of cases
\combpar{k} outperforms the baseline is significative (\eg{}, $117$ of
the \numFaults\ faults for $k=10$). One
of the following reasons justifies the baseline outperforming
\combpar{k}: (1)
a fault of omission or (2) the test has design limitations which \ds~is
not capable of handling (mentioned in Section~\ref{sec:slicing}).

@@

\begin{figure}[h]
		\centering
		\includegraphics[width=0.415\textwidth]{figures/performance.pdf}
		\vspace{-0.4cm}
		\caption{Delta Cost of Diagnosis ($\Delta{}C$) per $k$}
		\label{fig:diagnosis}
\end{figure}

% \Mar{Sofia, vc. precisa revisar com mais cuidado os dois proximos paragrafos...}

Figure~\ref{fig:boxplot} shows the distributions of $\Delta{}C$ considering all the 
cases where the predicate $\Delta{}C>0$ holds, \ie{}, it additionally excludes from the dataset
the cases of faults where the techniques perform equally for at least one $k$. 
To sum, although we observed a relatively small number of cases where $\Delta{}C$
is negative (\numOutPerformed) and a high number of cases $\Delta{}C$
is zero (\numEquallyPerformed), we conclude that there is a
considerable amount of cases where the use of \combpar{k} would be
very beneficial.
% On the left-hand-side, we show distributions of
% $\Delta{}C$, for different values of $k$, considering all cases where
% the predicate $\Delta{}C\geq{}0$ holds, \ie{}, we excluded the
% \numOutPerformed\ cases where \sfl{}$^{k}$ outperformed \combpar{k}
% (see Figure~\ref{fig:diagnosis}).

% The boxplots on the right-hand-side
% of the figure considers all the cases where the predicate
% $\Delta{}C>0$ holds, \ie{}, it additionally excludes from the dataset
% the cases of faults where the techniques perform equally.\Sof{Needs to be reverted}
% \Mar{sofia,
%   are you sure there are no zeros in this dataset? The min. in this
%   distibution (as per the base of the plot) seems to be 0.}\Sof{Yes, check hangouts}

\begin{figure}[h]
		\vspace{-0.2cm}
		\centering
		\includegraphics[width=0.45\textwidth]{figures/boxplot.pdf}
		\vspace{-0.4cm}

		\caption{Distributions of $\Delta{}C$ considering all cases where \combpar{k}
        outperformed the baseline}
		\label{fig:boxplot}
		\vspace{-0.2cm}
\end{figure}

In the following, we present statistical analysis of the techniques.
Table~\ref{table:st} reports descriptive statistics about the
distributions of $C$ for the techniques under different values of $k$.
In bold, we present the results of the $227$ faults that do not
comprise the cases that are result of the limitations mentioned before
(test design and fault of omission). The number in parentheses are
relative to the dataset including all data faults; \Sof{improve}note that results
for this dataset are \emph{not} encouraging.

@@

\begin{table}[h]
	\tiny
	\centering
	\setlength{\tabcolsep}{3pt}
	\begin{tabular}{@{}c|c|c|c|c@{}}
     \toprule
  & \multicolumn{2}{c|}{$k=5$} & \multicolumn{2}{c}{$k=10$} \\
  & SFL  & \comb\       & SFL                & \comb\              \\ \midrule
Mean  & \textbf{779.3} (584.3) & \textbf{738.2} (679.7)   &  \textbf{630.6} (516.4)    & \textbf{611.9} (614.0)   \\ \midrule
Median & \textbf{187.7} (49.4) & \textbf{167.2} (69.3)      & \textbf{167.2} (53.0) & \textbf{107.5} (56.9)\\ \midrule
Variance & \textbf{1218.3} (1223.1) & \textbf{1186.3} (1352.4)  &  \textbf{1058.4} (1148.1) &  \textbf{1048.6} (1294.9) \\ \midrule
Shapiro-Wilk &\makecell{$W$ = 0.67 \\ $p$ = 9.64 x $10^{-15}$} & \makecell{$W$ = 0.66 \\ $p$ = 5.59 x $10^{-15}$} & \makecell{$W$ = 0.63 \\ $p$ = 1.09 x $10^{-15}$} &  \makecell{$W$ = 0.63 \\ $p$ = 8.95 x $10^{-16}$}  \\ \midrule
Friedman & \multicolumn{4}{c}{\makecell{$\chi^{2}$ = 175.18 \\ p-value = 9.71 x $10^{-38}$}} \\
\bottomrule
\end{tabular}
  \caption {Statistical tests for $C$}
  \label{table:st}
\end{table}

@@

\begin{figure}[h]
	\vspace{-1cm}
		\includegraphics[width=0.45\textwidth]{figures/heatmap_nemenyi_result.pdf}
		\caption{Siegel post-hoc analysis results}
		\label{fig:performance}
		\vspace{-0.3cm}
\end{figure}

As the results are statistically significant, it is possible to state
that \combpar{k} has a positive effect on the localization of some
types of bugs. There is indeed a problem on using slicing on fault of
omission bugs that needs further research.  Although, this dataset has
more than $50\%$ of fault of omissions, our simple slicing approach
was capable of improving the ranking of $95$ faults for $k=5$ and
$115$ faults for $k=10$. These may be promissing results from what Automated
Program Repair (APR) techniques may directly benefit to improve the accuracy
of their own techniques.
